---
title: "GLM Birth Weights"
author: "Amit yaron"
date: "10/29/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
'
install.packages("ggcorrplot")
install.packages("ggplot2")
install.packages("cowplot")
install.packages("gridExtra")
install.packages("grid")
nstall.packages("ggpubr")
install.packages("lmtest")

'

```


#Preliminary Analysis of birth weights:

Upload the data 
```{r}
data <- read.table(file="birthwt.data.139.txt", header=TRUE )


low<-as.numeric(data[,1])
age<-as.numeric(data[,2])
lwt<-as.double(data[,3])
race<-as.character(data[,4])
smoke<-as.numeric(data[,5])
ht<-as.numeric(data[,6])
ui<-as.numeric(data[,7])
ftv<-as.numeric(data[,8])
ptl<-as.numeric(data[,9])
bwt<-as.numeric(data[,10])


```


change race to indicators : white or Other,Black
Create Matrix of continuous variables: her name Udata 
```{r}
white<- as.numeric(race == "white")
Udata<-cbind(low,age,lwt,white,smoke,ht,ui,ftv,ptl,bwt)
```



Histograms of continuous variables 
```{r}
par(mfrow=c(1,3))
hist(age,main = "Histogram of age", xlab="mother's age",border = "blue",col="green", xlim = c(10,45),breaks = 10,prob=TRUE)
lines(density(age))
hist(bwt,main = "Histogram of birth weight", xlab="birth weight",border = "blue",cex=1,col="red", xlim = c(500,5000),breaks = 10,prob=TRUE)
lines(density(bwt))
hist(lwt,main = "Histogram of mother's weight", xlab="mother's weight",border = "red",col="blue", xlim = c(80,250),breaks = 10,prob=TRUE)
axis(1,at =seq(80,250,by=120),labels = seq(80,250,by=120))
lines(density(lwt))


```

Histograms of continuous variables with -----ggplot2-------
```{r}
library(ggplot2)
library(cowplot)
par(mfrow=c(1,3))
age<-unlist(age)
age<- as.data.frame(age)
p1<-ggplot(data=age,aes(age)) + 
 geom_histogram(position="identity",aes(y =..density..),breaks=seq(10, 50, by = 2), 
                 col="red", 
                 fill="green", 
                 alpha = .2)+
   geom_density(aes(x=age,y=..density..))+
  labs(title="Histogram for Age", x="Age", y="Count")


bwt<-unlist(bwt)
bwt<- as.data.frame(bwt)
p2<-ggplot(data=bwt,aes(bwt)) + 
 geom_histogram(position="identity",aes(y =..density..),breaks=seq(1000, 5000, by = 200), 
                 col="red", 
                 fill="blue", 
                 alpha = .2)+
   geom_density(aes(x=bwt,y=..density..))+
  labs(title="Histogram of birth weight ", x="birth weight", y="Count")


lwt<-unlist(lwt)
lwt<- as.data.frame(lwt)
p3<-ggplot(data=lwt,aes(lwt)) + 
 geom_histogram(position="identity",aes(y =..density..),breaks=seq(80, 250, by = 10), 
                 col="red", 
                 fill="yellow", 
                 alpha = .2)+
   geom_density(aes(x=lwt,y=..density..))+
  labs(title="Histogram of mother weight ", x="mother weight", y="Count")



plot_grid(p1, p2,p3, labels = "AUTO")

```



Summery  of the variables with corrlsion matrix 
```{r}
library(ggcorrplot)
summary(Udata, digits = 1)
print("---------------")
table(low)
M<- cbind(lwt,bwt,age)
corr<-round(cor(M), 2)

ggcorrplot(corr, hc.order = TRUE, outline.col = "white", type = "lower", lab = TRUE)

```


Histograms of the indicator variables with table 
```{r}

par(mfrow=c(2,3))

library(ggplot2)
ggplot(data.frame(ftv), aes(x=ftv)) +
  geom_bar(fill="red",color="blue",)

ggplot(data.frame(ht), aes(x=ht)) +
  geom_bar(fill="blue",color="red")

ggplot(data.frame(low), aes(x=low)) +
  geom_bar(fill="green",color="red")

ggplot(data.frame(ptl), aes(x=ptl)) +
  geom_bar(fill="white",color="red")

ggplot(data.frame(smoke), aes(x=smoke)) +
  geom_bar(fill="black",color="red")

ggplot(data.frame(ui), aes(x=ui)) +
  geom_bar(fill="pink",color="red")

ggplot(data.frame(white),aes(x=white)) +
  geom_bar(fill="yellow",color="red")


ggplot(data.frame(race),aes(x=race)) +
  geom_bar(fill="orange",color="blue")


table(race)
print("--------------------------------------")
table(ftv)
print("--------------------------------------")
table(ht)
print("--------------------------------------")
table(low)
print("--------------------------------------")
table(ui)
print("--------------------------------------")
table(ptl)
print("--------------------------------------")
table(white)
print("--------------------------------------")
table(smoke)



```



Joint distribution of continuous variables:
```{r}
 library(GGally)
ggscatmat(cbind(age,bwt,lwt),alpha=0.8) +
   ggtitle("Joint distribution of continuous variables scatterplot matrix") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 2))

pairs(cbind(age,bwt,lwt))
   
```



Joint distribution of indicator variables:
```{r}
table(white,smoke)
prop.table(table(white,smoke))
print("----------------------------")
table(white,low)
prop.table(table(white,low))
print("----------------------------")
table(low,smoke)
prop.table(table(low,smoke))
print("----------------------------")
table(low,ht)
prop.table(table(low,ht))
print("----------------------------")
table(low,ui)
prop.table(table(low,ui))
print("----------------------------")
table(low,ptl)
prop.table(table(low,ptl))
```


Box-plot about the relation between the continuous variables to the category variables
```{r}

'
#low as age,bwt,lwt
par(mfrow=c(1,3))
boxplot(age ~low, xlab="low birth weight less than 2.5kg ", ylab="mother's age")
#boxplot(bwt ~low, xlab="low", ylab="bwt")
#axis(4, at=seq(0,5000, 500))
boxplot(lwt ~low, xlab="low", ylab="lwt")
axis(4, at=seq(0, 200, 10))


#age Box_plot! 
par(mfrow=c(2,3))
boxplot(age ~smoke, xlab="smoke", ylab="age")
boxplot(age ~ht, xlab="ht", ylab="age")
boxplot(age ~ui, xab="ui", ylab="age")
boxplot(age ~ftv, xab="ftv", ylab="age")
boxplot(age ~race, xab="race", ylab="age")
boxplot(age ~ptl, xab="ptl", ylab="age")

#bwt Box_plot! 
par(mfrow=c(2,3))
boxplot(bwt ~smoke, xlab="smoke", ylab="bwt")
boxplot(bwt ~ht, xlab="ht", ylab="bwt")
boxplot(bwt ~ui, xab="ui", ylab="bwt")
boxplot(bwt ~ftv, xab="ftv", ylab="bwt")
boxplot(bwt ~race, xab="race", ylab="bwt")
boxplot(bwt ~ptl, xab="ptl", ylab="bwt")

#lwt Box_plot! 
par(mfrow=c(2,3))
boxplot(lwt ~smoke, xlab="smoke", ylab="lwt")
boxplot(lwt ~ht, xlab="ht", ylab="lwt")
boxplot(lwt ~ui, xab="ui", ylab="lwt")
boxplot(lwt ~ftv, xab="ftv", ylab="lwt")
boxplot(lwt ~race, xab="race", ylab="lwt")
boxplot(lwt ~ptl, xab="ptl", ylab="lwt")

'


```



Box-plot about the relation between the continuous variables to the category variables with----ggplot2-----
```{r}
library(ggplot2)
library(ggpubr)
#-----------------------------------------------------age---------------------------------------------------------------------------------
#Plot of mother's age as baby birth weight
#

t1<- as.data.frame(cbind(age,low))

p1 <- ggplot(t1, aes(x=factor(low,labels =c("0","1") ), y=age ,color=factor(low,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="Baby birth weight less than 2.5kg", y = "Mother's age",color="low")
#
#Plot of mother's age as Smoking
#

t2<- as.data.frame(cbind(age,smoke))

p2 <- ggplot(t2, aes(x=factor(smoke,labels =c("0","1") ), y=age ,color=factor(smoke,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="Smoking status during pregnancy", y = "Mother's age",color="smoke")
#
#Plot of mother's age as white
#

t3<- as.data.frame(cbind(age,white))

p3 <- ggplot(t3, aes(x=factor(white,labels =c("0","1") ), y=age ,color=factor(white,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="Race white", y = "Mother's age",color="white")


ggarrange(p1,p2,p3+ rremove("x.text"), 
          ncol = 3, nrow = 1)

#-----------------------------------------------------age----------------------------------------------------------------------------
#
#Plot of mother's age as presence of uterine irritability
#

t4<- as.data.frame(cbind(age,ui))

p4 <- ggplot(t4, aes(x=factor(ui,labels =c("0","1") ), y=age ,color=factor(ui,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="presence of uterine irritability", y = "Mother's age",color="ui")

#
#Plot of mother's age as number of physician visits during the first trimester
#

t5<- as.data.frame(cbind(age,ftv))

p5 <- ggplot(t5, aes(x=factor(ftv,labels =c("0","1") ), y=age ,color=factor(ftv,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x=" physician visits during the first trimester", y = "Mother's age",color="ftv")

#
#Plot of mother's age as number of previous premature labours
#

t6<- as.data.frame(cbind(age,ptl))

p6 <- ggplot(t6, aes(x=factor(ptl,labels =c("0","1") ), y=age ,color=factor(ptl,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="previous premature labours", y = "Mother's age",color="ptl")


ggarrange(p4,p5,p6+ rremove("x.text"), 
          ncol = 3, nrow = 1)


#-----------------------------------------------lwt--------------------------------------------------------------------------------------
#
#Plot of mother's weight in pounds at last menstrual as low baby birth weight
#

t7<- as.data.frame(cbind(lwt,low))

p7 <- ggplot(t7, aes(x=factor(low,labels =c("0","1") ), y=lwt ,fill=factor(low,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="baby birth weight", y = "mother's weight",fill="low")

#
#Plot of mother's weight in pounds at last menstrual as smoke
#

t8<- as.data.frame(cbind(lwt,smoke))

p8 <- ggplot(t8, aes(x=factor(smoke,labels =c("0","1") ), y=lwt ,fill=factor(smoke,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="smoking status during pregnancy", y = "mother's weight",fill="smoke")

#
#Plot of mother's weight in pounds at last menstrual as race
#

t9<- as.data.frame(cbind(lwt,white))

p9 <- ggplot(t9, aes(x=factor(white,labels =c("0","1") ), y=lwt ,fill=factor(white,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="Race", y = "mother's weight",fill="white")


ggarrange(p7,p8,p9+ rremove("x.text"), 
          ncol = 3, nrow = 1)

#-----------------------------------------------lwt------------------------------------------------------------------------------------------

#
#Plot of mother's weight in pounds as history of hypertension ui
#

t10<- as.data.frame(cbind(lwt,ui))

p10 <- ggplot(t10, aes(x=factor(ui,labels =c("0","1") ), y=lwt ,fill=factor(ui,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="history of hypertension", y = "mother's weight",fill="ui")

#
#Plot of mother's weight in pounds as number of physician visits during the first trimester
#


t11<- as.data.frame(cbind(lwt,ftv))

p11 <- ggplot(t11, aes(x=factor(ftv,labels =c("0","1") ), y=lwt ,fill=factor(ftv,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="physician visits during the first trimester", y = "mother's weight",fill="ftv")

#
#Plot of mother's weight in pounds as number of previous premature labours
#


t12<- as.data.frame(cbind(lwt,ptl))

p12 <- ggplot(t12, aes(x=factor(ptl,labels =c("0","1") ), y=lwt ,fill=factor(ptl,labels = c("0","1")))) + 
  geom_boxplot()+
  labs(x="previous premature labours", y = "mother's weight",fill="ptl")




ggarrange(p10,p11,p12+ rremove("x.text"), 
          ncol = 3, nrow = 1)



#-----------------------------------------------bwt------------------------------------------------------------------------------------------

#
#Plot of birth weight in grams  as indicator of birth weight less than 2.5kg
#


t13<- as.data.frame(cbind(bwt,low))

p13 <- ggplot(t13, aes(x=factor(low,labels =c("0","1") ), y=bwt ,fill=factor(low,labels = c("0","1")))) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Dark2")+
  labs(x="birth weight", y = "birth weight in grams",fill="low")


#
#Plot of birth weight in grams  as smoking during pregnancy
#


t14<- as.data.frame(cbind(bwt,smoke))

p14 <- ggplot(t14, aes(x=factor(smoke,labels =c("0","1") ), y=bwt ,fill=factor(smoke,labels = c("0","1")))) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Dark2")+
  labs(x="smoking during pregnancy", y = "birth weight in grams",fill="smoke")

#
#Plot of birth weight in grams  as race white
#


t15<- as.data.frame(cbind(bwt,white))

p15 <- ggplot(t15, aes(x=factor(white,labels =c("0","1") ), y=bwt ,fill=factor(white,labels = c("0","1")))) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Dark2")+
  labs(x="Race", y = "birth weight in grams",fill="white")

ggarrange(p13,p14,p15+ rremove("x.text"), 
          ncol = 3, nrow = 1)
#-----------------------------------------------bwt------------------------------------------------------------------------------------------


#
#Plot of birth weight in grams  as presence of uterine irritability
#

t16<- as.data.frame(cbind(bwt,ui))

p16 <- ggplot(t16, aes(x=factor(ui,labels =c("0","1") ), y=bwt ,fill=factor(ui,labels = c("0","1")))) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Dark2")+
  labs(x="uterine irritability", y = "birth weight in grams",fill="ui")

#
#Plot of birth weight in grams  as physician visits during the first trimester
#

t17<- as.data.frame(cbind(bwt,ftv))

p17 <- ggplot(t17, aes(x=factor(ftv,labels =c("0","1") ), y=bwt ,fill=factor(ftv,labels = c("0","1")))) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Dark2")+
  labs(x="physician visits during the first trimester", y = "birth weight in grams",fill="ftv")

#
#Plot of birth weight in grams  as number of previous premature labours
#

t18<- as.data.frame(cbind(bwt,ptl))

p18 <- ggplot(t18, aes(x=factor(ptl,labels =c("0","1") ), y=bwt ,fill=factor(ptl,labels = c("0","1")))) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Dark2")+
  labs(x="previous premature labours", y = "birth weight in grams",fill="ptl")

ggarrange(p16,p17,p18+ rremove("x.text"), 
          ncol = 3, nrow = 1)

```




Fit OLS simple linear regression model
```{r}
lm_ols <- lm(bwt ~age+lwt+white+smoke+ht+ui+ftv+ptl,data=data)
summary(lm_ols)
plot(lm_ols$residuals, pch = 16, col = "red")

```


THe Best model logistic  on AIC ---Automatically select variables using two methods - Forward Selection and Backward Selection----
```{r}

glm_model<-glm(low~age+lwt+white+smoke+ht+ui+ftv+ptl,data=data, family = binomial(link = "logit"))

summary(glm_model)



best_model<-step(glm_model,direction = "both",trace=1)

summary(best_model)






```



THe Best model probit   on AIC  ---Automatically select variables using two methods - Forward Selection and Backward Selection----
```{r}
pro_model<-glm(low~age+lwt+white+smoke+ht+ui+ftv+ptl,data=data, family = binomial(link = "probit"))
 
summary(pro_model) 

pro_best_model<-step(pro_model,direction = "both",trace=1)

#significant variables verables are :   low ~ lwt + white + smoke + ht + ptl

summary(pro_best_model) 



```



Interaction and transformation on probit model 
```{r}

testmodel<-glm(low ~ lwt+white+smoke+(white*smoke) +ht+ptl,data=data, family = binomial(link = "probit"))
AIC(testmodel)

testmode2<-glm(low ~ lwt+white+smoke+ht+(smoke*ht)+ptl,data=data, family = binomial(link = "probit"))
AIC(testmode2)

testmode3<-glm(low ~ lwt+white+smoke+ht+(white*ht)+ptl,data=data, family = binomial(link = "probit"))
AIC(testmode3)

testmode4<-glm(low ~ lwt+white+smoke+(white*smoke) +ht+ptl,data=data, family = binomial(link = "probit"))
AIC(testmode4)

#Use log in the lwt !! lower the AIC


tpro_model<-glm(low ~ log(lwt)+white+smoke+ht+ptl,data=data, family = binomial(link = "probit"))
 
summary(tpro_model) 
```


 Test the model : Likelihood ratio test,BIC,check the overdispersion
```{r}
#Likelihood ratio test 
library(lmtest)
lrtest( tpro_model,pro_model)
#BIC 
BIC( tpro_model)
BIC(pro_model)
#check the overdispersion 1.050 almost 1 it good !! 
dp = sum(residuals(tpro_model,type ="pearson")^2)/tpro_model$df.residual
dp
```




Outliers and added-variable plot in r
```{r}
library(car)
#devlance response : response and pearson 
par(mfrow=c(1,2))
predicted_values = tpro_model$fitted	# estimated mu

response.resid1 = residuals(tpro_model, type="response")#
pearson.resid2= residuals(tpro_model, type="pearson")
summary(response.resid1)
summary(pearson.resid2)


plot(predicted_values,response.resid1)
identify(predicted_values,response.resid1)
plot(predicted_values,pearson.resid2)
identify(predicted_values,pearson.resid2)

par(mfrow=c(1,2))

deviance.resid3= residuals(tpro_model, type="deviance")
plot(predicted_values,deviance.resid3)
identify(predicted_values,deviance.resid3)
summary(deviance.resid3)

#cook distance
h.ii = hatvalues(tpro_model)
D.ii = cooks.distance(tpro_model)
plot(h.ii, D.ii)

summary(D.ii)
identify(h.ii, D.ii)

crPlots(tpro_model, one.page=TRUE)
avPlots(tpro_model, one.page=TRUE)

```



The Best model without Outliers 

```{r}
no_outliers=1:142
no_outliers=( no_outliers!=110 &  no_outliers!=117 & no_outliers!=128)
finel_tpro_model<-glm(low ~ log(lwt)+white+smoke+ht+ptl,data=data[no_outliers,], family = binomial(link = "probit"))
summary(finel_tpro_model)
dp2 = sum(residuals(finel_tpro_model,type ="pearson")^2)/tpro_model$df.residual
dp2

```




